#Lab_04 {lab_04}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here

# Load janitor here



```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.

setwd("C:/Users/benic/OneDrive/Documents/GitHub/Benicio_jour472/lab_04")
md_race <- read.csv("lab_04.csv")

# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)

```

## Answer questions

Q1: Name the five cities that had the largest Hispanic population as a percentage of the entire population in 2020 and call it top5_hispanic. Produce a top 5 table for each, selecting the place, 2020 Hispanic, 2020 total and your new calculation.

A1:
```{r}

# Write your code in this codeblock.
top5_hispanic <- md_race %>%
  select(place, x2020_hispanic, x2020_total) %>%
  mutate(top5_hispanic = x2020_hispanic / x2020_total * 100) %>%
  arrange(desc(top5_hispanic)) %>%
  head(5)
  


```

Q2: Compare the top 10 cities with the largest Black population as a percentage of the overall population in 2010 and 2020. Filter for places with more than 20,000 population in 2010. Call your results pct_black. Compare results for 2020 vs 2010. What is different? What surprised you? 

A2:
Looking at the data, Suitland, Lochearn, Landover, Baltimore and Fort Washington were the only places to see an increase in population. What surprised me is that, at least when looking at this data through the filter, the percentage increase was greater than the percentage decrease, even though Baltimore county has seen an overall decrease in its black population.


```{r}
#write your code here
pct_black <- md_race %>%
  select(place, x2020_black, x2020_total, x2010_black, x2010_total) %>%
  filter(x2010_total > 20000) %>%
  mutate(top10black2020 = x2020_black / x2020_total * 100,
         top10black2010 = x2010_black / x2010_total * 100) %>%
  arrange(desc(top10black2020)) %>%
  head(10)
```

Q3: Find the 15 communities that had the great gains in Hispanic population from 2010 to 2020 as measured by percentage change. Also calculate percentage change for the community's overall growth during the same period. Filter your results for communities with more than 10,000 people as of 2020. Your results will be hisp_growth. Look at your results, craft a one-paragraph story pitch for your editor. 

A3:

Story Pitch:
  In the past 10 years, the Hispanic population in Baltimore county has seen considerable growth. North Laurel, which had less than 200 Hispanic people 10 years ago, saw explosive growth across all of its demographics, with its Hispanic population reaching over 2,000 in 2020, and its overall population growing from 4000 to 25,000. The story would be interviewing local Hispanic families and people and seeing when they moved to North Laurel, and how they feel about the communities large and rapid growth. Are these families/people in the same economic bracket? Compared to the people who lived there in 2010, has the community become gentrified in any way?

```{r}
hisp_growth <- md_race %>%
  select(place, x2020_hispanic, x2020_total, x2010_hispanic, x2010_total) %>%
  filter(x2020_total > 10000) %>%
  mutate(hisp_change = (x2020_hispanic - x2010_hispanic)/x2010_hispanic * 100,
         total_change = (x2020_total - x2010_total)/x2010_total * 100) %>%
  arrange(desc(hisp_change)) %>%
  head(15)


```
