#Lab_03 {lab_03}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here

# Load janitor here


```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)

# your answer here
view(lab_03)
```


## Answer questions

Q1: Which five cities have the largest overall Hispanic population in 2020? Name the five cities that had the largest Hispanic population in 2020. 

A1:
```{r}

# Write your code in this codeblock.
lab_03 %>%
  select(place, x2020_hispanic) %>%
  slice_max(x2020_hispanic, n = 5)
   
# your answer here
Baltimore City at 32,627, Wheaton at 22787, see code for numbers, rest are Germantown, Silver Spring and Aspen Hill in descending order.
```

Q2: Find the cities with the largest Black population and arrange the results in descending order. 

A2:

```{r}
#write your code here
lab_03 %>%
  select(place, x2020_black) %>%
  slice_max(x2020_black, n = 5)
# your answer here
Baltimore City at 375,002, Waldorf at 47,903, the rest in descending order are Bowie, Clinton and Columbia.
```

Q3: Create a table called top_hispanic that shows the top 15 communities with the largest Hispanic population in 2020, and also display the 2010 Hispanic population and the total community population for 2020. Look at your results - what stands out to you? Craft a one-paragraph story pitch for your editor. 

A3:

```{r}

# your answer here
lab_03 %>%
  select(place, x2020_hispanic, x2010_hispanic, x2020_total, x2010_total) %>%
  slice_max(x2020_hispanic, n = 15) %>%
  slice_max(x2010_hispanic, n = 15)

   In the past 10 years, the Hispanic population of Maryland has seen steady overall growth. In places like Langley Park, Chillu and Adelphi, the demographic makes up at least 50% of the population, if not more. Even in Baltimore, which has seen a steady population decline in the past decade, the Hispanic population has only grown. The pitch would be to interview Hispanic families from these areas and ask them how the Hispanic community has been in the past decade. If they were immigrants, was moving to a predominantly Hispanic area a priority? If it wasn't, and they happened to be there by chance, has it been an overall positive experience?

```
