---
editor: 
  markdown: 
    wrap: sentence
editor_options: 
  chunk_output_type: inline
---

#Lab_05 {lab_05}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**).
\* modify existing code as directed (look for **Task**).
\* write code in empty codeblocks provided to answer questions included (look for **Q**).
\* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
library(tidyverse)
library(janitor)
```

```{r}
# Turn off scientific notation
options(scipen=999)

```

### Load Life Expectancy Data

You'll need to load :

CDC_Life_Census_Tract_2010_2015.csv

It is included in your lab_05.zip file.

```{r}
read.csv("CDC_Life_Census_Tract_2010_2015.csv")
```

About the data: Life Expectancy at Birth for U.S.
States and [Census Tracts, 2010-2015](https://www.cdc.gov/nchs/data-visualization/life-expectancy/index.html)

It shows estimates of U.S. life expectancy at birth by state and census tract for the period 2010-2015.
Estimates were produced for 67,148 census tracts, covering all states and the District of Columbia (D.C.). These estimates come from the National Center for Health Statistics (NCHS), the National Association for Public Health Statistics and Information Systems (NAPHSIS), and the Robert Wood Johnson Foundation (RWJF).

**Task** Create a codeblock below this task, then load the data sets and assign them to appropriate variable names.

## Answer questions

**Q1.** In your lab_05 folder, there is a csv called CDC_Life_Census_Tract_2010_2015.csv.
It contains all of the life expectancy data for the United States.

You are going to examine this data and learn the range of death rates in Baltimore and then in Maryland in general.

Do the following: - Load tidyverse and janitor - Load the data CDC_Life_Census_Tract_2010_2015.csv into a dataframe called us_death.
Include the code to show the number of records.
Clean the names using Janitor - Filter this according to Maryland into second table called md_death - Filter this according to Baltimore City into a third table called balt_death

**A1.**

```{r}
us_death <- read.csv("CDC_Life_Census_Tract_2010_2015.csv") %>%
  clean_names() %>%
  print(nrow(us_death)) 

```

```{r}
  md_death <- us_death %>%
  filter(state == "Maryland")
  print(md_death)
```

```{r}
balt_death <- us_death %>%
  filter(county == "Baltimore city, MD")
  print(balt_death)
```

**Q2.** For this question, load a modified dataset called balt_death_census.csv.

Call it balt_death2

This has the same information but with the Baltimore neighborhoods.

Produce summary statistics showing the data distribution by quartile, median, mean, miniumum and maximum values on the life expectancy rate.

Show the neighborhoods with a life expectancy rate below the 1st quartile, arranged in ascending order.

**A2.**

```{r}
balt_death2 <- read.csv("balt_death_census.csv")
  summary(balt_death2) %>%
```
```{r}
below <- balt_death2 %>%
  filter(life_expectancy < 69.58) %>%
  arrange(life_expectancy)
```



**Q3.** Use md_death, the Maryland life expectancy rates.
Produce summary statistics of the state of Maryland life expectancy rate.
Produce a table with the 20 communities with the lowest life expectancy.
Produce a table with the 20 communities with the highest life expectancy.

Then use us_death, and determine the average life expectancy nationwide.

Then write 150-200 words on this document below your code that compares your findings of Baltimore, the state and nation.

**A3.**

```{r}
md_death %>%
  summary(md_death)
```
```{r}
top_20_life_expectancy <- md_death %>%
  arrange(desc(life_expectancy)) %>%
  head(20) %>%
  select(state, county, life_expectancy)
```
```{r}
bottom_20_life_expectancy <- md_death %>%
  arrange(life_expectancy) %>%
  head(20) %>%
  select(state, county, life_expectancy)
```
```{r}
summary(us_death)
```
When comparing the average life expectancy of Baltimore, Maryland and the U.S. as a whole, Baltimore comes in at around 73, Maryland at around 79, and the U.S at around 78. I'm not surprised to see Baltimore below both the state and country average, as the factors of poverty and crime would lower that statistic. I'm also not surprised to see Maryland above the country average, as there are far less values to take into account that would skew the number lower. Also, Maryland is a relatively affluent state, so the quality of living goes up along with that.

Looking at Baltimore more closely, I'm surprised that its minimum life expectancy value is higher than that of the country, with the former being nearly 63, and the latter being nearly 57. I cannot think of a logical reason for this, but if I had to guess, maybe regions where there's a lot of obesity/illness are skewing the national average down due to their impact on life expectancy.

**Q4.**

Use balt_death.
Use mutate and the case_when command.

Categorize all Baltimore City Census tracts as being above or below the city's average life expectancy rate.

Then produce a table that summarizes the number of Census tracts above or below the average life expectancy rate.
Write two sentences summarizing your findings.

Pro Tip: See this chapter section for a refresher on case_when: https://wellsdata.github.io/data_journalism_class/\_book/r-mutating.html#a-more-powerful-use

```{r}
balt_death_mutate <- balt_death %>%
  mutate(
    average_life_expectancy = case_when(
      life_expectancy >= 73.04 ~ "above",
      life_expectancy <= 73.04 ~ "below"
    )
  )

balt_death_mutate %>%
  select(state, county, life_expectancy, average_life_expectancy) %>%
  arrange(average_life_expectancy)

```
Based on these findings, 87 census tracts are above the average life expectancy of 73.04.On the other hand, 93 census tracts fall below the average. Therefore, the majority of Baltimore census tracts with appliable data fall below the average.