#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse
```{r}
library(tidyverse)
library(janitor)
```


## Load data

**Task**: Load four tables of U.S. Census data.  Each table has one row per Maryland county with 2020 population information.  Each table has these same two columns:

```{r}
maryland_black_data <- read.csv("data/maryland_black.csv")
maryland_hispanic_data <- read.csv("data/maryland_hispanic.csv")
maryland_white_data <- read.csv("data/maryland_white.csv")
maryland_population_data <- read.csv("data/maryland_population.csv")
```


* geoid - a unique identifier for each county
* name - the county name

The third column in each table represents the total population in one of four demographic categories.

* maryland_population.csv - total_pop - total population 
* maryland_hispanic.csv - hispanic_pop - total Hispanic population
* maryland_white.csv - white_pop - total White alone population
* maryland_black.csv - black_pop - total Black alone population

Some of the information in the tables may need to be cleaned or validated in order to produce the correct answer.  

## Questions

**Q1**  What Maryland county has the largest Hispanic population (by percentage of total population)? 

```{r}
fixed_maryland_hispanic_data <- maryland_hispanic_data %>%
  mutate(name = case_when(
    name == "Montgomery County,Maryland" ~ "Montgomery County, Maryland",
    TRUE ~ name
  ))
```


```{r}
maryland_hispanic_joined <- fixed_maryland_hispanic_data %>%
  inner_join(maryland_population_data, by = "geoid", "name") %>%
  mutate(maryland_hispanic_joined, percent_of_pop = hispanic_pop / total_pop) %>%
  print(mutated_hispanic_pop)
  
```

**A1** 
Montgomery County has the largest Hispanic population by percentage of total population at nearly 20%.

**Q2**  What is the total population of people who identify as white alone in Maryland? 
```{r}
fixed_maryland_white_data <- maryland_white_data %>%
  mutate(white_pop = case_when(
    name == "Somerset County, Maryland" ~ 13541,
    TRUE ~ white_pop
  ))
sum(fixed_maryland_white_data$white_pop)
```

**A2**  
3275079 people identify as white alone in Maryland.

**Q3** What is the average Black population in all Maryland counties? 
```{r}
fixed_maryland_black_data <- maryland_black_data %>%
  mutate(black_pop = case_when(
    name == "Baltimore County, Maryland"  ~ 243698,
    name == "Baltimore city, Maryland" ~ 367387,
    TRUE ~ black_pop
  ))

mean(fixed_maryland_black_data$black_pop)
```

**A3** 
The average Black population in Maryland counties is 74838.